\documentclass[11pt]{article}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{bm}
\usepackage{a4}
\usepackage[top=2.5cm,bottom=2.5cm,left=2.5cm,right=2.5cm]{geometry}
\usepackage[usenames, dvipsnames]{xcolor}
\usepackage{url}
\usepackage{hyperref}

\setcitestyle{aysep={}}

\definecolor{lincolngreen}{rgb}{0.11, 0.35, 0.02}

\hypersetup{
  colorlinks,%
  linkcolor=lincolngreen,%
  citecolor=black,%
  filecolor=black,%
  urlcolor=lincolngreen
}

\makeatletter
\newcommand{\rmnum}[1]{\romannumeral #1}
\newcommand{\Rmnum}[1]{\expandafter\@slowromancap\romannumeral #1@}
\makeatother

\newcommand{\euler}{e}
\newcommand{\ud}{\,\mathrm{d}}


\begin{document}
\newtheorem%
{pro}{Proposition}
\newtheorem%
{lem}{Lemma}
\newtheorem%
{cor}{Corollary}


\begin{center}
Universit\'e de Gen\`eve \\
Global Studies Institute \\
\vspace{6mm}
\huge{J2P216 SE: International Cooperation and Conflict} \\
\vspace{6mm}
\large{Spring 2016}
\end{center}


\vspace{8mm}
\noindent Time: Thu/Fri, 12:15-14:00 \\
Place: M 2150 (Thu) / M S050 (Fri) \\
Insctructor: Reto W\"uest \\
Email: \href{mailto:reto.wuest@unige.ch}{reto.wuest@unige.ch} \\
Office Hours: Wed, 13:00-15:00, SIP A7$^{\prime}$ \\
Course Website: \href{http://retowuest.github.io/j2p216/}{http://retowuest.github.io/j2p216/}


\section*{Description}

This seminar is an introduction to the analysis of international cooperation and conflict. The seminar begins with a brief review of the primary theoretical paradigms in the field and the levels at which states interact. It then addresses the question of why states have common interests, providing them with an incentive to cooperate. Finally, we examine why states sometimes fail to cooperate. In particular, we explore how uncertainty, commitment problems, and mobilization can lead to bargaining failure.


\section*{Resources}


\subsection*{Readings}


\subsection*{Course Website}


\subsection*{Piazza}


\section*{Grading}


\newpage
\section*{Schedule}

\vspace{2mm}
\subsection*{Introduction to International Relations}

~\\
\textsc{March 3/4: Introduction} \\

\noindent \underline{Required Readings}

\begin{itemize}
\item Walt, Stephen M. 1998. \href{http://www.jstor.org/stable/1149275}{``International Relations: One World, Many Theories.''} \textit{Foreign Policy} 110: 29-46.
\end{itemize}

~\\
\noindent \textsc{March 10/11: Levels of Analysis in International Relations} \\

\noindent \underline{Required Readings}

\begin{itemize}
\item Singer, J. David. 1961. \href{http://www.jstor.org/stable/2009557}{``The Level-of-Analysis Problem in International Relations.''} \textit{World Politics} 14(1): 77-92.
\end{itemize}

\noindent \underline{Further Readings}

\begin{itemize}
\item Waltz, Kenneth N. 1959. \textit{Man, the State, and War: A Theoretical Analysis.} New York: Columbia University Press. [Chapters 2, 4, 6.]
\end{itemize}


\vspace{2mm}
\subsection*{Why Do States Bargain?}

~\\
\textsc{March 17/18: Commerce} \\

\noindent \underline{Required Readings}

\begin{itemize}
\item \textbf{[P]} Gowa, Joanne and Edward D. Mansfield. 1993. \href{http://www.jstor.org/stable/2939050}{``Power Politics and International Trade.''} \textit{American Political Science Review} 87(2): 408-420.
\item \textbf{[P]} Gartzke, Erik. 2007. \href{http://onlinelibrary.wiley.com/doi/10.1111/j.1540-5907.2007.00244.x/abstract}{``The Capitalist Peace.''} \textit{American Journal of Political Science} 51(1): 166-191.
\end{itemize}

\noindent \underline{Further Readings}

\begin{itemize}
\item
\end{itemize}


\noindent \textsc{March 24/25: Territory} \\

\noindent \underline{Required Readings}

\begin{itemize}
\item Walter, Barbara F. 2003. \href{http://onlinelibrary.wiley.com/doi/10.1111/j.1079-1760.2003.00504012.x/abstract}{``Explaining the Intractability of Territorial Conflict.''} \textit{International Studies Review} 5(4): 137-153.
\item Simmons, Beth A. 2005. \href{http://www.jstor.org/stable/30045138}{``Rules over Real Estate: Trade, Territorial Conflict, and International Borders as Institution.''} \textit{Journal of Conflict Resolution} 49(6): 823-848.
\item Branch, Jordan. 2011. \href{http://www.jstor.org/stable/23016102}{``Mapping the Sovereign State: Technology, Authority, and Systemic Change.''} \textit{International Organization} 65(1): 1-36.
\item Carter, David B. and H. E. Goemans. 2011. \href{http://www.jstor.org/stable/23016813}{``The Making of the Territorial Order: New Borders and the Emergence of Interstate Conflict.''} \textit{International Organization} 65(2): 275-309.
\end{itemize}

\noindent \underline{Further Readings}

\begin{itemize}
\item
\end{itemize}


\noindent \textsc{April 7/8: Ideology} \\

\noindent \underline{Required Readings}

\begin{itemize}
\item Lake, David A. 1992. \href{http://www.jstor.org/stable/1964013}{``Powerful Pacifists: Democratic States and War.''} \textit{American Political Science Review} 86(1): 24-37.
\item Bukovansky, Mlada. 1999. \href{http://www.jstor.org/stable/20097590}{``The Altered State and the State of Nature: The French Revolution and International Politics.''} \textit{Review of International Studies} 25(2): 197-216.
\end{itemize}

\noindent \underline{Further Readings}

\begin{itemize}
\item
\end{itemize}


\noindent \textsc{April 14/15: Identity} \\

\noindent \underline{Required Readings}

\begin{itemize}
\item Huntington, Samuel P. 1993. \href{http://www.jstor.org/stable/20045621}{``The Clash of Civilizations?''} \textit{Foreign Affairs} 72(3): 22-49.
\item Henderson, Errol A. and Richard Tucker. 2001. \href{http://www.jstor.org/stable/3096113}{``Clear and Present Strangers: The Clash of Civilizations and International Conflict.''} \textit{International Studies Quarterly} 45(2): 317-338.
\end{itemize}

\noindent \underline{Further Readings}

\begin{itemize}
\item Posen, Barry R. 1993. \href{http://www.tandfonline.com/doi/abs/10.1080/00396339308442672?journalCode=tsur20}{``The Security Dilemma and Ethnic Conflict.''} \textit{Survival} 35(1): 27-47.
\item Gagnon, V. P., Jr. 1994/95. \href{http://www.jstor.org/stable/2539081}{``Ethnic Nationalism and International Conflict: The Case of Serbia.''} \textit{International Security} 19(3): 130-166.
\item Snyder, Jack and Karen Ballentine. 1996. \href{http://www.jstor.org/stable/2539069}{``Nationalism and the Marketplace of Ideas.''} \textit{International Security} 21(2): 5-40.
\end{itemize}


\vspace{2mm}
\subsection*{Why Does Bargain Fail?}

~\\
\noindent \textsc{April 21/22: Bargaining Theory} \\

\noindent \underline{Required Readings}

\begin{itemize}
\item Powell, Robert. 2002. \href{http://search.ebscohost.com/login.aspx?direct=true&db=buh&AN=6790109&site=ehost-live}{``Bargaining Theory and International Conflict.''} \textit{Annual Review of Political Science} 5: 1-30. [Read carefully: pp. 1-14, 19-23; skim: pp. 14-19, 23-27]
\end{itemize}

\noindent \underline{Further Readings}

\begin{itemize}
\item Muthoo, Abhinay. 2000. \href{http://isites.harvard.edu/fs/docs/icb.topic1188138.files/Week%203/Muthoo_2000.pdf}{``A Non-Technical Introduction to Bargaining Theory.''} \textit{World Economics} 1(2): 145-166.
\item Reiter, Dan. 2003. \href{http://www.jstor.org/stable/3687811}{``Exploring the Bargaining Model of War.''} \textit{Perspectives on Politics} 1(1): 27-43.
\end{itemize}

~\\
\noindent \textsc{April 28/29: Uncertainty} \\

\noindent \underline{Required Readings}

\begin{itemize}
\item Fearon, James D. 1995. \href{http://www.jstor.org/stable/2706903}{``Rationalist Explanations for War.''} \textit{International Organization} 49(3): 379-414.
\item Levy, Jack S. 1983. \href{http://www.jstor.org/stable/2010176}{``Misperception and the Causes of War: Theoretical Linkages and Analytical Problems.''} \textit{World Politics} 36(1): 76-99.
\end{itemize}

\noindent \underline{Further Readings}

\begin{itemize}
\item Gartzke, Erik. 1999. \href{http://www.jstor.org/stable/2601290}{``War Is in the Error Term.''} \textit{International Organization} 53(3): 567-587.
\item Garfinkel, Michelle R. and Stergios Skaperdas. 2000. \href{http://jcr.sagepub.com/content/44/6/793.short}{``Conflict without Misperceptions or Incomplete Information: How the Future Matters.''} \textit{Journal of Conflict Resolution} 44(6): 793-807.
\item Levento\u{g}lu, Bahar and Ahmer Tarar. 2008. \href{http://www.jstor.org/stable/29734250}{``Does Private Information Lead to Delay or War in Crisis Bargaining?''} \textit{International Studies Quarterly} 52(3): 533-553.
\end{itemize}

~\\
\noindent \textsc{May 5/6: Commitment Problems} \\

\noindent \underline{Required Readings}

\begin{itemize}
\item Powell, Robert. 2006. \href{http://www.jstor.org/stable/3877871}{``War as a Commitment Problem.''} \textit{International Organization} 60(1): 169-203.
\end{itemize}

\noindent \underline{Further Readings}

\begin{itemize}
\item Walter, Barbara F. 1997. \href{http://www.jstor.org/stable/2703607}{``The Critical Barrier to Civil War Settlement.''} \textit{International Organization} 51(3): 335-364.
\item Lake, David A. 2010/11. \href{http://www.mitpressjournals.org/doi/abs/10.1162/ISEC_a_00029#.VorZipMrLdQ}{``Two Cheers for Bargaining Theory: Assessing Rationalist Explanations of the Iraq War.''} \textit{International Security} 35(3): 7-52.
\item Findley, Michael G. and Joseph K. Young. 2011. \href{http://onlinelibrary.wiley.com/doi/10.1111/j.1468-2478.2011.00647.x/full}{``Terrorism, Democracy, and Credible Commitments.''} \textit{International Studies Quarterly} 55(2): 357-378.
\end{itemize}

~\\
\noindent \textsc{May 12/13: Mobilization} \\

\noindent \underline{Required Readings}

\begin{itemize}
\item Fearon, James D. 1997. \href{http://jcr.sagepub.com/content/41/1/68.short}{``Signaling Foreign Policy Interests: Tying Hands versus Sinking Costs.''} \textit{Journal of Conflict Resolution} 41(1): 68-90.
\item Slantchev, Branislav L. 2005. \href{http://www.jstor.org/stable/30038963}{``Military Coercion in Interstate Crises.''} \textit{American Political Science Review} 99(4): 533-547.
\end{itemize}

\noindent \underline{Further Readings}

\begin{itemize}
\item Debs, Alexandre and Nuno P. Monteiro. 2014. \href{http://dx.doi.org/10.1017/S0020818313000192}{``Known Unknowns: Power Shifts, Uncertainty, and War.''} \textit{International Organization} 68(1): 1-31.
\end{itemize}

~\\
\noindent \textsc{May 19/20: Domestic Politics} \\

\noindent \underline{Required Readings}

\begin{itemize}
\item Putnam, Robert D. 1988. \href{http://www.jstor.org/stable/2706785}{``Diplomacy and Domestic Politics: The Logic of Two-Level Games.''} \textit{International Organization} 42(3): 427-460.
\item Fearon, James D. 1994. \href{http://www.jstor.org/stable/2944796}{``Domestic Political Audiences and the Escalation of International Disputes.''} \textit{American Political Science Review} 88(3): 577-592.
\end{itemize}

\noindent \underline{Further Readings}

\begin{itemize}
\item Schultz, Kenneth A. 1998. \href{http://www.jstor.org/stable/2586306}{``Domestic Opposition and Signaling in International Crises.''} \textit{American Political Science Review} 92(4): 829-844.
\item Chiozza, Giacomo and H. E. Goemans. 2004. \href{http://www.jstor.org/stable/1519919}{``International Conflict and the Tenure of Leaders: Is War Still `Ex Post' Inefficient?''} \textit{American Journal of Political Science} 48(3): 604-619.
\item Jackson, Matthew O. and Massimo Morelli. 2007. \href{http://www.jstor.org/stable/30034096}{``Political Bias and War.''} \textit{American Economic Review} 97(4): 1353-1373.
\item Weeks, Jessica L. 2008. \href{http://www.jstor.org/stable/40071874}{``Autocratic Audience Costs: Regime Type and Signaling Resolve.''} \textit{International Organization} 62(1): 35-64.
\end{itemize}


\vspace{2mm}
\subsection*{Conclusion}

~\\
\noindent \textsc{May 26/27: Conclusion} \\


\end{document}
